Improving Software Maintainability through automatic refactoring of Code Clones

### 2019-11-06 saner/paper.tex
Updates from Overleaf

\begin{abstract}
Duplication in source code is often seen as one of the most harmful types of technical debt as it increases the size of the codebase and creates implicit dependencies between fragments of code. To remove such anti-patterns, the codebase should be refactored. Many tools aid in the de©tection process of such duplication problems but do not determine whether refactoring an identified clone would improve the maintainability.

We address this shortcoming by first analyzing what preconditions apply to be able to refactor clones. We then propose a tool to detect clones, analyze their context and automatically refactor a subset of them. We use a set of established metrics to determine the impact of the applied refactorings on the maintainability of the system. Based on these results, one could decide which clones would improve system design if refactored. We evaluate our approach over a large corpus of open-source Java projects.

We identify the factors of interest characterizing the impact on maintainability of automatically applied refactorings. We find that the biggest influencing factor is token volume. We define token volume as the combined number of tokens in all clone instances in a clone class. The majority of duplicates with a token volume of 63 or more tokens improve maintainability when refactored. The amount of external data that needs to be passed to the merged location of the duplicate is the other important factor: the majority of clones requiring more than two external parameters decrease maintainability.
\end{abstract}

@@

\section{Introduction}
Duplicate fragments in source code (also named ``code clones'') are often seen as one of the most harmful types of technical debt~\cite{fowler2018refactoring}. Duplicate fragments create implicit dependencies that make the code harder to maintain as the resolution of erroneous behavior in one location may have to be applied to all the cloned code as well~\cite{ostberg2014automatically}. Apart from that, code clones can contribute up to 25\% of the code size~\cite{bruntink2005use}.

%Clone detection tools often use thresholds to determine which clones should be considered. Most clone detection tools use the number of lines \cite{kamiya2002ccfinder, svajlenko2016bigcloneeval}, statements \cite{higo2013revisiting, kamalpriya2017enhancing} and/or tokens \cite{ragkhitwetsagul2019siamese, jiang2007deckard} as thresholds to determine whether clones should be considered. We outline the shortcomings of these thresholds to find clones such that the maintainability of the source code is improved when refactored, and propose new thresholds to control this characteristic.
% Current clone detection techniques are limited to measuring the size of clones to determine whether they should be considered. Because of this, the output of clone detection tools is often of limited assistance to the developer in the refactoring process.
To reduce duplication of a codebase, there are several refactoring techniques that can be applied. The most used refactoring technique to deal with code clones is ``Extract Method'' \cite{fowler2018refactoring}. This technique entails that the duplicated code is moved to a new method, which is then called from all locations where the duplicated code was used. To be able to execute such a refactoring, several preconditions apply \cite{tsantalis2015assessing}.
%\todo[inline]{still missing link on how current solutions do not help improving maintainability - our message from monday night is not crystal clear yet}
In this study, we analyze which preconditions apply to determine whether a clone could and should be automatically refactored. We list criteria that influence the maintainability of clones when refactored. We evaluate the context of clones to determine which refactoring techniques are best suited. We define the context of clones as the inheritance relation between clone instances in a clone class and the location of clone instances in the codebase. We showcase the proposed technique as a tool that automatically refactors clones by extracting new methods out of duplicated code.

@@

Most clone detection tools can be configured using thresholds. These thresholds indicate the minimum number of lines, tokens and/or statements that must be spanned for duplicate fragments to be considered clones. Often, such thresholds are intuitively chosen~\cite{li2006cp, roy2009mutation} or based on a quartile distribution of empirical data~\cite{alves2010deriving}. Using the maintainability score we can find support for which thresholds should be chosen to increase the chance to find clones that improve maintainability when refactored.% \todo{Check whether this really fits here}

@@

This paper is organized as follows. In Sec.~\ref{sec:background} we revisit key clone refactoring background. Sec.~\ref{sec:relatedwork} discusses state-of-the-art in automated clone refactoring research. In Sec.~\ref{sec:research} we propose a definition for clones such that they could and should be automatically refactored. In Sec.~\ref{sec:clonerefactor} we introduce our clone refactoring tool. In Sec.~\ref{sec:experimentalsetup} we explain the setup of our experiments; we show the results in Sec.~\ref{sec:results} and discuss them in Sec.~\ref{sec:discussion}.
%We evaluate our approach by comparing the maintainability of before- and after snapshots of a codebase for each duplication problem that is refactored. We define which factors have the greatest influence on whether a clone improves maintainability when refactored. We find a set of thresholds by which clones can be detected that are more likely to improve maintainability when refactored. This allows for more a accurate suggestion of code clones for refactoring and can assist in the refactoring process.

@@

\subsection{Thresholds}
CR has refactored 12.710 clone classes and measured the change in the selected metrics (Sec.~\ref{sec:metrics}). %\todo{use the formulas labels to refer here.
Using the presented formulas (Sec.~\ref{sec:metricformula}) we determine how the characteristics of the extracted method (see Sec.~\ref{sec:characteristics}) influence the maintainability of the resulting codebase after refactoring. We explore the data obtained by comparing the before- and after snapshots of the system for each separate refactoring. Using this data, we find supporting evidence regarding which thresholds are most likely to find clones that should be refactored to improve maintainability.

@@

\subsubsection{Return Value}
Table~\ref{tab:return} shows how the return value of the extracted method influences the maintainability of the resulting system. The return categories are ordered by their scores. %\todo{this sentence about relation impact seems more suitable for discussion}
Scores do not deviate much (-0.18 to 0.19), indicating that the return category has a minor impact on maintainability. When the result of the call to the extracted method is directly returned, the maintainability score is the highest. When no value is returned, the maintainability score is the lowest. %\todo{this sentence about relation impact seems more suitable for discussion}
The main reason that no return value is lowest, is that it is linked to a higher number of parameters required for the extracted method.

@@

%\todo{Redo plot, explain axes units}
\begin{figure}[H]
  \includegraphics[width=1\columnwidth]{img/arguments}
  \caption{Influence of number of method parameters on system maintainability.}
  \label{fig:arguments}
\end{figure}

@@

We measured the change in four maintainability metrics
%before- and after applying each refactoring 
to determine the impact of each refactoring on system maintainability. We found that the most prominent factor influencing maintainability is the size of the clone. We found that the threshold lies at a clone volume of 63  %\todo{wasn't this 63?}
tokens for all clone instances combined for system maintainability to increase after refactoring the clone. Another factor with a major impact on maintainability is the number of parameters that the extracted method requires to get all required data. We noticed that the inheritance relation of the clone and the return value of the extracted method has only a minor impact on system maintainability.

>>>>

\begin{abstract}
Duplication in source code is often seen as one of the most harmful types of technical debt as it increases the size of the codebase and creates implicit dependencies between fragments of code. To remove such anti-patterns, the codebase should be refactored. Many tools aid in the detection process of such duplication problems but do not determine whether refactoring an identified clone would improve the maintainability. Moreover, manual code clone refactoring is often tedious and could benefit from an automated tool.

We address this shortcoming by first analyzing what preconditions apply to be able to automatically refactor clones. We then propose a tool to detect clones, analyze their context and automatically refactor a subset of them. We use a set of established metrics to determine the impact of the applied refactorings on the maintainability of the system. Based on these results, one could decide which clones would improve system design if refactored. We evaluate our approach over a large corpus of open-source Java projects.

We identify the factors of interest characterizing the impact on the maintainability of automatically applied refactorings. We find that the biggest influencing factor is token volume. We define token volume as the combined number of tokens in all clone instances in a clone class. The majority of duplicates with a token volume of 63 or more tokens improve maintainability when refactored. The amount of external data that needs to be passed to the merged location of the duplicate is the other important factor: the majority of clones requiring more than two external parameters decrease maintainability.
\end{abstract}

@@

\section{Introduction}
Duplicate fragments in source code (also named ``code clones'') are often seen as one of the most harmful types of technical debt~\cite{fowler2018refactoring}. Duplicate fragments create implicit dependencies that make the code harder to maintain as the resolution of erroneous behavior in one location may have to be applied to all the cloned code as well~\cite{ostberg2014automatically}. Apart from that, code clones can contribute up to 25\% of the code size~\cite{bruntink2005use}. However, manual code clone refactoring is often tedious and could benefit from an automated tool~\cite{mazinanian2016jdeodorant}.

%Clone detection tools often use thresholds to determine which clones should be considered. Most clone detection tools use the number of lines \cite{kamiya2002ccfinder, svajlenko2016bigcloneeval}, statements \cite{higo2013revisiting, kamalpriya2017enhancing} and/or tokens \cite{ragkhitwetsagul2019siamese, jiang2007deckard} as thresholds to determine whether clones should be considered. We outline the shortcomings of these thresholds to find clones such that the maintainability of the source code is improved when refactored, and propose new thresholds to control this characteristic.
% Current clone detection techniques are limited to measuring the size of clones to determine whether they should be considered. Because of this, the output of clone detection tools is often of limited assistance to the developer in the refactoring process.
To reduce duplication of a codebase, there are several refactoring techniques that can be applied in an automated manner. The most used refactoring technique to deal with code clones is ``Extract Method'' \cite{fowler2018refactoring}. This technique entails that the duplicated code is moved to a new method, which is then called from all locations where the duplicated code was used. To be able to execute such a refactoring, several preconditions apply \cite{tsantalis2015assessing}.
%\todo[inline]{still missing link on how current solutions do not help improving maintainability - our message from monday night is not crystal clear yet}
In this study, we analyze which preconditions apply to determine whether a clone could and should be automatically refactored. We list criteria that influence the maintainability of clones when refactored. We evaluate the context of clones to determine which refactoring techniques are best suited. We define the context of clones as the inheritance relation between clone instances in a clone class and the location of clone instances in the codebase. We showcase the proposed technique as a tool that automatically refactors clones by extracting new methods out of duplicated code.

@@

Most clone detection tools can be configured using thresholds. These thresholds indicate the minimum number of lines, tokens and/or statements that must be spanned for duplicate fragments to be considered clones. Often, such thresholds are intuitively chosen~\cite{li2006cp, roy2009mutation} or based on a quantile distribution of empirical data~\cite{alves2010deriving}. Using the maintainability score we can find support for which thresholds should be chosen to increase the chance to find clones that improve maintainability when refactored.% \todo{Check whether this really fits here}

@@

This paper is organized as follows. In Sec.~\ref{sec:background} we revisit key clone refactoring background. Sec.~\ref{sec:relatedwork} discusses state-of-the-art in automatic clone refactoring research. In Sec.~\ref{sec:research} we propose a definition for clones such that they could and should be automatically refactored. In Sec.~\ref{sec:clonerefactor} we introduce our clone refactoring tool. In Sec.~\ref{sec:experimentalsetup} we explain the setup of our experiments; we show the results in Sec.~\ref{sec:results} and discuss them in Sec.~\ref{sec:discussion}.
%We evaluate our approach by comparing the maintainability of before- and after snapshots of a codebase for each duplication problem that is refactored. We define which factors have the greatest influence on whether a clone improves maintainability when refactored. We find a set of thresholds by which clones can be detected that are more likely to improve maintainability when refactored. This allows for more a accurate suggestion of code clones for refactoring and can assist in the refactoring process.

@@

\subsection{Thresholds}
CR has refactored 12.710 clone classes and measured the change in the selected metrics (Sec.~\ref{sec:metrics}). %\todo{use the formulas labels to refer here.
Using the formulas in Eq.~\ref{eq:scoredev} and~\ref{eq:scoreref}
(Sec.~\ref{sec:metricformula}) we determine how the characteristics of the extracted method (see Sec.~\ref{sec:characteristics}) influence the maintainability of the resulting codebase after refactoring. We explore the data obtained by comparing the before- and after snapshots of the system for each separate refactoring. Using this data, we find supporting evidence regarding which thresholds are most likely to find clones that should be refactored to improve maintainability.

@@

\subsubsection{Return Value}
Table~\ref{tab:return} shows how the return value of the extracted method influences the maintainability of the resulting system. The return categories are ordered by their scores. %\todo{this sentence about relation impact seems more suitable for discussion}
Scores do not deviate much (-0.18 to 0.19), indicating that the return category has a minor impact on maintainability. When the result of the call to the extracted method is directly returned, the maintainability score is the highest. When no value is returned, the maintainability score is the lowest. %\todo{this sentence about relation impact seems more suitable for discussion}
The main reason that "Void" is lowest, is that it is linked to a higher number of parameters required for the extracted method.

@@

%\todo{Redo plot, explain axes units}
\begin{figure*}
  \includegraphics[width=1\textwidth]{img/argu}
  \caption{Influence of number of method parameters on system maintainability.}
  \label{fig:arguments}
\end{figure*}

@@

We measured the change in four maintainability metrics
%before- and after applying each refactoring 
to determine the impact of each refactoring on system maintainability. We found that the most prominent factor influencing maintainability is the size of the clone. We found that the threshold lies at a clone volume of 63  %\todo{wasn't this 63?}
tokens for all clone instances combined for system maintainability to increase after refactoring the clone. Another factor with a major impact on maintainability is the number of parameters that the extracted method requires to get all required data: when a clone is refactored to a method that requires more than 2 parameters it is more likely to decrease maintainability than to increase it. We noticed that the inheritance relation of the clone and the return value of the extracted method has only a minor impact on system maintainability.
